\documentclass[8pt,landscape]{extarticle}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsfonts,bm}
\usepackage{multicol}
\usepackage{geometry}
\geometry{left=0.8cm,right=0.8cm,top=0.8cm,bottom=0.8cm,landscape}
\usepackage{parskip}
\usepackage{enumitem}
\setlist{leftmargin=*,nosep,itemsep=0pt,topsep=0pt}
\pagestyle{empty}
\setlength{\columnsep}{0.3cm}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.2ex}

\newcommand{\R}{\mathbb{R}}
\newcommand{\norm}[1]{\|#1\|}
\newcommand{\1}{\mathbf{1}}
\newcommand{\0}{\mathbf{0}}
\newcommand{\diag}{\text{diag}}

\begin{document}

\begin{multicols*}{3}

% ============================================================
\subsection*{1. Vectors \& Basic Ops}
$e_i^Ta=a_i$\\
Inner product: $x^Ty = \sum x_i y_i$, $\approx 2n$ flops. \\
Norms: $\norm{x}_2 = \sqrt{x^Tx}$, $\norm{x}_1 = \sum|x_i|$, $\norm{x}_\infty = \max|x_i|$ \\
Angle: $x^Ty = \norm{x}\norm{y}\cos\theta$ \\
Cauchy-Schwarz: $|x^Ty| \le \norm{x}\norm{y}$\\
Linear independence: $\sum\alpha_i v_i=0 \Rightarrow \alpha_i=0$ \\
Basis: $n$ independent vectors in $\R^n$ \\
Orthonormal: $v_i^Tv_j=0$ for $i\ne j$, $\|v_i\|=1$; $n$ orthonormal vectors in $\R^n$ form ON basis.\\
Affine function: $f(x)=a^Tx+b$ (linear + constant).

\subsection*{2. Matrices \& Linear Systems}
Positive Definite: $x^TAx>0\ \forall x\ne0$. \\
Positive Semidefinite: $x^TAx\ge0\ \forall x$; \\
$A^{-1}$ exists iff rows/cols lin. indep. iff rows/cols form basis \\
Solve $Ax=b$: Diagonal $O(n)$, Triangular $O(n^2)$, Permutation $O(0)$ \\
Solving Linear Systems: \\
LU: $A=PLU$, factor $\frac23n^3$, solve $2n^2$. \\
Cholesky: $A=LL^T$ for symmetric PD, factor $\frac13n^3$, solve $2n^2$.

\subsection*{3. Least Squares (L3)}
Tall $A\in\R^{m\times n}$, $m>n$: usually no exact $Ax=b$. \\ 
LS problem: $\min_x \|Ax-b\|_2^2$, residual $r=Ax-b$. \\
Column view: LS finds $Ax^*\in\operatorname{col}(A)$ closest to $b$ \\
Gradient: $\nabla\|Ax-b\|_2^2=2A^T(Ax-b)$. \\
Optimality: $A^T(Ax^*-b)=0$ $\Rightarrow$ residual $r^*\perp\operatorname{col}(A)$. \\
Solve: $A^TAx^*=A^Tb$. \\
If $A$ full col rank: $x^*=(A^TA)^{-1}A^Tb$ \\
Gram $G=A^TA$: symmetric PSD always; PD/invertible iff $A$ has indep. cols. \\
Solve via Cholesky: factor $A^TA=LL^T$, solve $LL^Tx=A^Tb$ \\
Complexity: form $A^TA$: $2mn^2$; Cholesky: $\frac13 n^3$; prefactored solve (new $b$): $\approx 2mn$ \\

\subsection*{4. Least Squares Data Fitting}
Data: $(x^{(i)},y^{(i)})$, $i=1,\ldots,N$. \\
Model:  $\hat f(x)=\theta_1f_1(x)+\cdots+\theta_pf_p(x)$, linear in $\theta$. \\
Matrix form: $A_{ij}=f_j(x^{(i)})$, $\hat y=A\theta$. \\
LS: $\min_\theta \|A\theta-y\|_2^2$, $A^TA\theta^*=A^Ty$. \\
Residual $r^{(i)}=y^{(i)}-\hat y^{(i)}$.
MSE $=\frac1N\sum_i(r^{(i)})^2=\frac1N\|A\theta-y\|_2^2$; RMSE $=\sqrt{\mathrm{MSE}}$. \\
Line: $\hat f(x)=\theta_1+\theta_2x$, $A=[\mathbf1\;\;x_d]$. \\
Polynomial: $f_j(x)=x^{j-1}$, $\hat f(x)=\sum_{j=1}^p\theta_jx^{j-1}$,
$A_{ij}=(x^{(i)})^{j-1}$ (Vandermonde). \\
AR($M$): $\hat z_{t+1}=\theta_1z_t+\cdots+\theta_Mz_{t-M+1}$;
$y^{(i)}=z_{M+i}$, $x^{(i)}=(z_{M+i-1},\ldots,z_i)$. \\
General data fitting = linear regression on transformed features:
$\tilde x=(f_2(x),\ldots,f_p(x))$, intercept $f_1(x)=1$. \\
Validation: train/test split; overfitting = low train MSE + high test MSE.
$k$-fold CV: train on $k-1$ folds, test on held-out fold; choose lowest test error, then retrain on all data.

\subsection*{5. Multi-Objective Least squares}
Problem: $\min_x \sum_{i=1}^k \lambda_i\|A_ix-b_i\|^2$, with $\lambda_i>0$. \\
Equivalent stacked LS:
$\min_x\left\|\begin{bmatrix}\sqrt{\lambda_1}A_1\\ \vdots\\ \sqrt{\lambda_k}A_k\end{bmatrix}x-
\begin{bmatrix}\sqrt{\lambda_1}b_1\\ \vdots\\ \sqrt{\lambda_k}b_k\end{bmatrix}\right\|^2$ \\
Solution: $(\tilde A^T\tilde A)x^* = \tilde A^T\tilde b$ \\
Sum of distances to points: $\sum_{i=1}^k\|x-a_i\|^2
= \left\|\begin{bmatrix}I\\ \vdots\\ I\end{bmatrix}x
-\begin{bmatrix}a_1\\ \vdots\\ a_k\end{bmatrix}\right\|^2$.
Here $A\in\R^{kn\times n}$ stacks $k$ copies of $I_n$, $b\in\R^{kn}$ stacks $a_i$.\\
Bi-criterion: $\min J_1+\lambda J_2$. If $J_2$ too big, increase $\lambda$; if $J_1$ too big, decrease $\lambda$. \\
Trade-off curve: $(J_1(x^*(\lambda)), J_2(x^*(\lambda)))$, Pareto optimal \\
Tikhonov: $\min \norm{Ax-y}^2 + \lambda\norm{x}^2 = \left\|\begin{bmatrix}A\\\sqrt{\lambda}I\end{bmatrix}x-\begin{bmatrix}y\\0\end{bmatrix}\right\|^2$ \\ 
Normal equation: $(A^TA+\lambda I)x=A^Ty$ \\
Smoothing: $\min \norm{Ax-y}^2 + \lambda\norm{Dx}^2$ ($D$ = diff matrix) \\
Normal equation:  $(A^TA+\lambda D^TD)x=A^Ty$ \\

\subsection*{6. Constrained Least Squares}
Problem: $\min \norm{Ax-b}^2$ subject to $Cx=d$ \\
Lagrangian $L(x,z)=\|Ax-b\|^2+z^T(Cx-d)$. \\
KKT: $\begin{bmatrix}2A^TA & C^T \\ C & 0\end{bmatrix}\begin{bmatrix}x\\z\end{bmatrix} = \begin{bmatrix}2A^Tb\\d\end{bmatrix}$ \\
KKT invertible iff $C$ has indep. rows and $[A;C]$ has indep cols \\
Least norm: $\min_x\|x\|^2$ s.t. $Cx=d$ = CLS with $A=I,b=0$. \\
Portfolio: weights $w$, $\mathbf1^Tw=1$; $w_i<0$ short; leverage $L=\|w\|_1$. \\
Return matrix $R\in\R^{T\times n}$, portfolio returns $r=Rw$.
$\operatorname{avg}(r)=\mathbf1^Tr/T$, 
$\operatorname{std}(r)=\|r-\operatorname{avg}(r)\mathbf1\|/\sqrt T$. \\
Annualized return $=P\operatorname{avg}(r)$; annualized risk $=\sqrt P\operatorname{std}(r)$. \\
Portfolio CLS:
$\min_w\|Rw-\rho\mathbf1\|^2$
s.t. $\mathbf1^Tw=1$, $\mu^Tw=\rho$, $\mu=R^T\mathbf1/T$. \\
KKT:
$\begin{bmatrix}
2R^TR & \mathbf1 & \mu\\
\mathbf1^T & 0 & 0\\
\mu^T & 0 & 0
\end{bmatrix}
\begin{bmatrix}w\\z_1\\z_2\end{bmatrix}
=
\begin{bmatrix}2\rho T\mu\\1\\\rho\end{bmatrix}$. \\
Two-fund theorem: $w^*(\rho)=w_0+\rho v$ affine in target return.

\subsection*{7. Linear Optimization / LP}
LP general form: $\min c^Tx$ subj. to $Ax\le b$, $Dx=f$; inequalities are computed elementwise. \\
Feasible $x$ satisfies constraints. $x^*$ optimal if feasible and $c^Tx^*\le c^Tx$ for all feasible $x$.
Optimal value $p^*=c^Tx^*$. \\
Special cases: infeasible $\Rightarrow p^*=+\infty$; unbounded below $\Rightarrow p^*=-\infty$; optimizer may be nonunique. \\
Feasibility problem (find any $x$ subj. to LP constraints): $\min 0$ subj. to constraints of the LP\\
Standard form: $\min c^Tx$ s.t. $Ax=b$, $x\ge0$. \\
Transformations to standard form: \\
maximize $c^Tx \Rightarrow$ minimize $-c^Tx$. \\
$Ax\le b \Rightarrow Ax+s=b,\ s\ge0$; \\ 
$Ax\ge b \Rightarrow Ax-s=b,\ s\ge0$. \\
$x_i\le0 \Rightarrow x_i=-y_i,\ y_i\ge0$; \\
$x_i$ free $\Rightarrow x_i=x_i^+-x_i^-$, $x_i^+,x_i^-\ge0$.

\subsection*{8. Piecewise Linear Optimization}
Convex PL: $f(x)=\max_i(a_i^Tx+b_i)$ \\
Minimization: $\min_x f(x)\equiv \min t$ s.t. $a_i^Tx+b_i\le t,\ \forall i$. \\
Sum of PL: use separate $t_j$ for each max term; minimize $\sum_j t_j$. \\
$\min\norm{Ax-b}_\infty$: $\min t$ s.t. $-t\1\le Ax-b\le t\1$ \\
$\min\norm{Ax-b}_1$: $\min \1^Tu$ s.t. $-u\le Ax-b\le u$ \\
$\ell_2^2$: sum squared errors. 
$\ell_1$: sum absolute errors $\sum_i |r_i|$.
$\ell_\infty$: max absolute error $\max_i |r_i|$.\\
General abs trick: $|y_i|\le u_i \iff y_i\le u_i,\ -y_i\le u_i$. \\
Constraint $\|x\|_1\le k$: introduce $u$, $-u\le x\le u$, $\mathbf1^Tu\le k$. \\
Sparse recovery: $\min_x\|x\|_1$ s.t. $Ax=y$
$\equiv \min_{x,u}\mathbf1^Tu$ s.t. $-u\le x\le u$, $Ax=y$; $\ell_1$ promotes sparsity. \\
SVM separable: find $a,b$ s.t. $s_i(a^Tv_i+b)\ge1$, $s_i\in\{-1,1\}$; LP feasibility. \\
Nonseparable hinge loss:
$\min_{a,b}\sum_i\max\{0,1-s_i(a^Tv_i+b)\}$
$\equiv \min\mathbf1^Tu$ s.t. $u_i\ge0$, $u_i\ge1-s_i(a^Tv_i+b)$.\\
Worst-case max of linear costs: $\min_x \max_k (p^k)^Tx
\equiv \min_{x,v} v$ s.t. $(p^k)^Tx\le v,\ \forall k$.\\
Smoothness / total variation: $\sum_{t=1}^{T-1}|x_{t+1}-x_t|$\\
LP via $y_t\ge x_{t+1}-x_t$, $y_t\ge -(x_{t+1}-x_t)$; minimize $\mathbf1^Ty$.


\subsection*{9. Geometry \& Polyhedra (L9)}
Hyperplane: $\{x\mid a^Tx=b\}$; halfspace: $\{x\mid a^Tx\le b\}$. 
Polyhedron: $P=\{x\mid Ax\le b\}$ = finite halfspaces \\
%{Convex combo: $\sum_i\alpha_ix_i$, $\alpha_i\ge0$, $\sum_i\alpha_i=1$.
%Convex hull $\operatorname{conv}\{x_i\}$ = all convex combos. Polyhedra are convex. \\}
Extreme point: cannot write $x=\alpha y+(1-\alpha)z$ with $y,z\in P$, $y,z\ne x$, $\alpha\in(0,1)$. \\
Vertex: unique minimizer of some linear objective over $P$. \\
Active constraints: $I(\bar x)=\{i\mid a_i^T\bar x=b_i\}$. \\
Standard form $P=\{x\mid Ax=b,\ x\ge0\}$, $A\in\R^{m\times n}$ full row rank.
Choose $m$ indep. columns $A_B$, set $x_N=0$, solve $A_Bx_B=b$.
If $x_B\ge0$, then $x$ is a Basic Feasible Sol (BFS). \\
Equivalence: vertex $\iff$ extreme point $\iff$ BFS. \\
Degenerate BFS: $>n$ active constraints (in standard form, some basic variable $=0$) \\
$P$ has an extreme point iff $P$ contains no line. Nonempty bounded polyhedron has a BFS.
If LP optimum exists and $P$ has an extreme point, then some optimum is an extreme point.

\subsection*{11. The Simplex Method}
Idea: move BFS $\to$ neighboring BFS with lower cost \\
Reduced cost: $\bar c_j = c_j - c_B^T A_B^{-1}A_j$; $\bar c_j=0$ for basic \\
Optimal: all $\bar c_j \ge 0$ \\
If $\bar c_j<0$, bringing $x_j$ into basis decreases cost \\
Feasible direction: $d_j=1$, solve $A_B d_B = -A_j$ \\
Step length: $\theta^* = \min_{\{i\in B\mid d_i<0\}} (-x_i/d_i)$ (exiting index) \\
Unbounded: if $d_B\ge0$, then $x+\theta d$ feasible $\forall\theta\ge0$ and cost $\to-\infty$. \\
Simplex iteration: (1) $A_B^Tp=c_B$, $\bar c = c-A^Tp$; (2) if $\bar c\ge0$ stop; (3) choose $j$ with $\bar c_j<0$; (4) $A_B d_B=-A_j$, $d_j=1$; (5) if $d_B\ge0$ unbounded; (6) $\theta^*$ from min ratio; (7) $x\leftarrow x+\theta^*d$; (8) update basis \\
Finite convergence (if non-degenerate): objective strictly decreases cost and there are finitely many BFSs\\
Degeneracy can cause cycling; Bland's rule (smallest index entering/leaving) prevents cycling


\subsection*{12. Simplex Implementation}
Need initial BFS for $\min c^Tx$ s.t. $Ax=b$, $x\ge0$. \\
Phase (I): assume $b\ge0$ wlog; if $b_i<0$, multiply row $i$ by $-1$. 
add artificial $y$, $\min \1^Ty$ s.t. $Ax+y=b$, $x\ge0,y\ge0$, start $x=0,y=b$; \\
if opt $>0$ original LP infeasible; if $=0$ get BFS, drop $y$, goto (II) \\
Phase (II): solve original from Phase I basis \\
Bland's rule: entering = smallest index with $\bar c_j<0$; leaving = smallest index attaining min ratio \\
Complexity: per iteration $O(n^2)$ (matrix update); worst-case $2^n$; average $O(n)$

\subsection*{13--14. Duality}
Standard primal: $\min c^Tx$ s.t. $Ax=b,\ x\ge0$
$\Rightarrow$ dual: $\max -b^Ty$ s.t. $A^Ty+c\ge0$, $y$ free. \\
Inequality primal: $\min c^Tx$ s.t. $Ax\le b$
$\Rightarrow$ dual: $\max -b^Ty$ s.t. $A^Ty+c=0,\ y\ge0$. \\
Mixed primal: $\min c^Tx$ s.t. $Ax\le b,\ Cx=d$
$\Rightarrow$ dual: $\max -b^Ty-d^Tz$ s.t. $A^Ty+C^Tz+c=0,\ y\ge0$, $z$ free. \\
Sign convention: write constraints as $Ax-b\le0$ or $Ax-b=0$. Inequality dual vars $\ge0$; equality dual vars free. \\
Weak duality: primal feasible $x$, dual feasible $y$ $\Rightarrow$ dual obj $\le$ primal obj; for inequality form $-b^Ty\le c^Tx$. \\
Duality gap: $c^Tx+b^Ty\ge0$; if gap $=0$, $x,y$ optimal. \\
Strong duality: if primal has optimum, dual has optimum and $p^*=d^*$. Primal unbounded $\Rightarrow$ dual infeasible; dual unbounded $\Rightarrow$ primal infeasible. \\
Complementary slackness: $y_i(b_i-a_i^Tx)=0$. Hence $y_i>0\Rightarrow a_i^Tx=b_i$; $a_i^Tx<b_i\Rightarrow y_i=0$. \\
KKT (checklist for proving a sol is opt.): primal feas ($Ax\le b$); dual feas ($A^Ty+c=0,\ y\ge0$); comp. slack ($y_i(Ax-b)_i=0\ \forall i$)\\
Farkas: exactly one holds: (i) $\exists x\ge0$ s.t. $Ax=b$; 
(ii) $\exists y$ s.t. $A^Ty\ge0,\ b^Ty<0$.

\subsection*{15. Sensitivity Analysis}
Add variable $x_{n+1}$: old $(x^*,0)$ still primal feasible.
still opt iff $A_{n+1}^Ty^*+c_{n+1}\ge0$; else add variable \& use primal simplex \\
Add constraint $a_{m+1}^Tx=b_{m+1}$: old dual $(y^*,0)$ still dual feasible. still optimal iff $a_{m+1}^Tx^*=b_{m+1}$; else add constraint \& use dual simplex. \\
RHS perturbation: $p^*(u)=\min\{c^Tx\mid Ax=b+u,\ x\ge0\}$ = optimal val when RHS $b$ is changed to $b+u$.
Global lower bound: $p^*(u)\ge p^*(0)-u^Ty^*$. \\
If basis $B$ unchanged:
\begin{itemize}
  \item $x_B^*(u)=A_B^{-1}(b+u)=x_B^*+A_B^{-1}u$, $x_N^*(u)=0$
  \item $p^*(u)=p^*(0)-u^Ty^*=-(b+u)^Ty^*$
  \item $\nabla p^*(0)=-y^* = $ shadow price; one more unit of RHS $i$ changes objective by $-y_i^*$
  \item Objective perturbation: if $c\to c+\epsilon d$ and basis unchanged,
$p^*(\epsilon)=p^*(0)+\epsilon d^Tx^*$.
\end{itemize}


\subsection*{16. Network Flow}
Directed network: $m$ nodes, $n$ arcs \\
Arc-node incidence $A\in\R^{m\times n}$: $A_{ij}=1$ if arc $j$ leaves $i$, $-1$ if enters, else $0$ \\
Min cost flow: $\min c^Tx$ s.t. $Ax=b$, $0\le x\le u$, $\1^Tb=0$ \\
Total unimodularity: node-arc incidence is TU \\
Integrality: if $A$ TU and $b,u$ integer, BFS are integer \\
Max flow: add arc sink$\to$source with cost $-1$, cap $\infty$; source $1$, sink $m$, $e=(1,0,\ldots,0,-1)$:
$\max t$ s.t. $Ax=te$, $0\le x\le u$. \\
As min-cost flow: $\min -t$ s.t. $[A\ -e]\binom{x}{t}=0$, $0\le x\le u$, $t\ge0$.\\
Shortest path from $1$ to $m$: $\min c^Tx$ s.t. $Ax=e$, $0\le x\le1$,
where $e=(1,0,\ldots,0,-1)$\\
By TU, LP solution is binary, so $x_j=1$ iff arc $j$ is in path.\\
Assignment: nodes = people \& tasks; arcs = possible assignments.
$x_k$ = flow on arc $k$ ($x_{ij}=1$ if person $i$ assigned to task $j$, else $0$). 
$A\in\R^{(\#nodes)\times(\#arcs)}$ incidence matrix:
$A_{ik}=1$ if arc $k$ leaves node $i$, $-1$ if arc $k$ enters node $i$, $0$ otherwise.
$b=(\1_N,-\1_N)$: people have supply $+1$, tasks have demand $-1$.
Solve $\min c^Tx$ s.t. $Ax=b$, $0\le x\le1$.


\subsection*{17--18. Interior-Point Methods}
Newton: solve $h(x)=0$, $x^{k+1}=x^k+\Delta x$, $Dh(x^k)\Delta x=-h(x^k)$ \\
For $\min c^Tx$ s.t. $Ax\le b$, add slack $s\ge0$: $Ax+s=b$. \\
LP optimality: $h=[Ax+s-b;\; A^Ty+c;\; SY\1]=0$, $s,y\ge0$ \\
Residuals: $r_p=Ax+s-b$, $r_d=A^Ty+c$, $\mu=s^Ty/m$ \\
Central path (smoothed): $s_iy_i=\tau$, $s,y>0$; follow as $\tau\to0$ \\
Newton system ($\sigma\in[0,1]$, $\sigma=1$ centering, $\sigma=0$ pure Newton):
\[
\begin{bmatrix}0&A&I\\ A^T&0&0\\ S&0&Y\end{bmatrix}
\begin{bmatrix}\Delta y\\\Delta x\\\Delta s\end{bmatrix}
= \begin{bmatrix}-r_p\\-r_d\\-SY\1+\sigma\mu\1\end{bmatrix}
\]
Primal-dual path-following: choose $\sigma$, solve system, line search for max $\alpha$ keeping $s,y>0$, update. \\
Barrier: $\min c^Tx-\tau\sum_i\log(s_i)$ s.t. $Ax+s=b$; barrier keeps $s_i>0$. \\
Stop when $\|r_p\|,\|r_d\|,\ s^Ty$ are small. \\
\textbf{Mehrotra predictor-corrector}: \\
1. Predictor ($\sigma=0$), get $\Delta_a$, compute $\alpha_p,\alpha_d,\mu_a$ \\
2. $\sigma = (\mu_a/\mu)^3$ \\
3. Corrector RHS $-SY\1 - \Delta S_a\Delta Y_a\1 + \sigma\mu\1$ \\
Complexity: IPM average $O(\log n)$ iterations, $O(n^3)$ per iter \\
\textbf{IPM vs Simplex}: IPM variants polynomial, hard to warm-start, good large/sparse; Simplex exponential worst-case, needs BFS/Phase I, can warm-start, good small/medium.


\subsection*{20--21. Integer Optimization}
Problem: $\min c^Tx$ s.t. $Ax\le b$, $x_i\in\mathbb Z$ for $i\in I$ \\
Binary: $x_i\in\{0,1\}$. LP relaxation removes integrality; for min, $p^*_{\rm rel}\le p^*_{\rm IP}$. \\
Logic: at most one $\mathbf1^Tx\le1$; $x_1$ only if $x_2$: $x_1\le x_2$; both/neither: $x_1=x_2$. \\
Knapsack: $\max c^Tx$ s.t. $a^Tx\le b$, $x_i\in\{0,1\}$. \\
If-then ($y=0\Rightarrow x=0$) via $0\le x\le My$ (Big-M) \\ 
Disjunction: $a^Tx\le b$ or $d^Tx\le f$ via $a^Tx\le b+yM$, $d^Tx\le f+(1-y)M$. \\
Cardinality: $\operatorname{card}(x)\le k$ via $\sum_i y_i\le k$, $-My_i\le x_i\le My_i$, $y_i\in\{0,1\}$. \\
Restricted values $x\in\{a_1,\ldots,a_d\}$: $x=a^Tz$, $\mathbf1^Tz=1$, $z\in\{0,1\}^d$. \\
Better formulation = tighter relaxation. If $P_{\rm rel,1}\subset P_{\rm rel,2}$, then $p^*_{\rm rel,2}\le p^*_{\rm rel,1}\le p^*_{\rm IP}$ for min. \\
Ideal formulation: LP relaxation gives integer optimum \\

\textbf{Branch-and-bound}: \\
- Bound: $L_j$ (relaxation lower bound), $U_j$ (heuristic upper bound) \\
- Global: $L=\min L_j$, $U=\min U_j$, gap $U-L$ \\
- Branch: if $\bar x_i$ fractional, split into $x_i\le\lfloor\bar x_i\rfloor$ and $x_i\ge\lceil\bar x_i\rceil$. \\
- Prune if relaxation infeasible, relaxation solution integer, or $L_j\ge U$ \\
Incumbent = best integer feasible solution found so far; its value gives global upper bound $U$\\


\subsection*{22--23. Robust Optimization / Uncertainty}
Uncertain constraint: $a^Tx\le b$ with $a\in U$ \\
Robust constr: $a^Tx\le b\ \forall a\in U
\iff \max_{a\in U}a^Tx\le b$. \\
Parametrize $a=\bar a+Mu$, $u\in U$. Let $w=M^Tx$.
Then robust constraint is $\bar a^Tx+\max_{u\in U}w^Tu\le b$. \\
\textbf{Box uncertainty} ($\|u\|_\infty\le\rho$):
$\max w^Tu=\rho\|w\|_1$\\
Robust: $\bar a^Tx+\rho\|M^Tx\|_1\le b$\\
LP: $\bar a^Tx+\rho\mathbf1^Tt\le b$, $-t\le M^Tx\le t$. \\
\textbf{$1$-norm uncertainty} ($\|u\|_1\le\rho$):
$\max w^Tu=\rho\|w\|_\infty$ \\
Robust: $\bar a^Tx+\rho\|M^Tx\|_\infty\le b$ \\
LP: $\bar a^Tx+\rho\tau\le b$, $-\tau\le(M^Tx)_i\le\tau$. \\
\textbf{Polyhedral uncertainty} ($U=\{u\mid Fu\le g\}$):
$\max_{Fu\le g}w^Tu$ dualizes to
$\min_\lambda g^T\lambda$ s.t. $F^T\lambda=w$, $\lambda\ge0$\\
Robust LP: $\bar a^Tx+g^T\lambda\le b$, $F^T\lambda=M^Tx$, $\lambda\ge0$. \\
Larger $\rho$ = more conservative. Robust solution trades nominal objective for better worst-case feasibility/performance.\\

\textbf{Scenario Approach}:\\
Given samples $d_1,\ldots,d_N$ of uncertain coefficient $a$,
replace robust constraint $a^Tx\le b\ \forall a\in U$ with sample constraints
$d_i^Tx\le b,\ i=1,\ldots,N$. \\
Scenario LP: $\min c^Tx$ s.t. $d_i^Tx\le b,\ i=1,\ldots,N$. 
Dataset itself becomes uncertainty set. \\
Sample bound: if $d_i$ i.i.d., choose 
$N\ge \frac{2}{\epsilon}\left(\ln\frac1\beta+n-1\right)$
to get violation probability $\le\epsilon$ with confidence $\ge1-\beta$. \\
$n$ = decision variables, $\epsilon$ = violation tolerance, $\beta$ = confidence failure probability.
Pros: distribution-free, simple, no reformulation. 
Cons: random solution, may violate unseen scenarios, larger $N$ increases solve time. \\


\end{multicols*}
\end{document}